Spatial Analysis

### 2022-09-19 Scripts/spatial_analisis.Rmd
included new figure showing where each state is within the US

# Visual exploration (o.k! )
# Visual exploration (o.k! )
ggplot() +
  geom_sf(data = st_as_sf(map("state", plot = FALSE, fill = TRUE))) +
  geom_sf(data = land_sf, aes(fill = state)) +
  scale_fill_manual(values = state_pallet) +
  my_ggtheme_m(leg_pos = "right")

@@

state_sf <-  st_read("~/OneDrive - UW-Madison/FederalAndStateWaters/FederalAndStateWaters.gdb") %>%
  janitor::clean_names() %>% 
  mutate(jurisdicti = str_to_lower(jurisdicti)) %>% 
  filter(jurisdicti %in% States) %>% 
  mutate(state = str_to_sentence(jurisdicti)) %>% 
  st_transform(crs = 4326) %>%   # for matching projections
  st_simplify(preserveTopology = TRUE, dTolerance = 10000) #0.1 for paper

@@

# quantiles_df <- overall_port_data %>% 
#   group_by(species_name) %>% 
#   summarise(
#     "95th" = quantile(year_ton, 0.95),
#     "12th" = quantile(year_ton, 0.75),
#     "50th" = quantile(year_ton, 0.50)) %>% 
#   gather("tresh","quant",`95th`:`50th`)
# 
# 
# top_portsb <- overall_port_data %>% 
#   left_join(quantiles_df,
#             by = "species_name") %>% 
#   mutate(
#     treshold = ifelse(year_ton >= quant,tresh,NA) 
#   )

@@

# Visual exploration of all together (o.k! )
ggplot() +
  geom_sf(data = eez_sf, aes()) +
  geom_sf(data = land_sf, aes(fill = state), alpha = 0.3) +
  geom_sf(data = state_sf, aes(color = state), fill = NA) +
  geom_point(data = subset(noaa_ports_complete, tresh == "12th"), aes(x = lon, y = lat,
                                                                      shape = species_name), size = 3, alpha = 0.7) +
  scale_color_manual(values = state_pallet) +
  scale_fill_manual(values = state_pallet) +
  theme_dark() +
  coord_sf(ylim = c(33,48))

@@

#Load top port data 
ports_df <- my_path("G","AcrossBoundaries/Data/Partial", "top_ports.csv", read = T) %>% 
  filter(tresh == "12th")

@@

# Ports data
ports_df <- my_path("G","AcrossBoundaries/Data/Partial", "top_ports.csv", read = T) %>% 
  filter(tresh == "12th") %>%
  group_by(species_name,state_postal) %>% 
  summarise(n =n()) %>% 
  select(-n)
# distinct(c(,state_postal), .keep_all = T) # No need for duplication

@@

length(state_pallet)

@@

@@

my_path("D","Partial","top_ports.csv", read = T) %>% 
  filter(tresh == "12th",
         state_postal == "NY") %>% 
  ggplot() +
  geom_point(
    aes(
      x = lon,
      y = lat,
      color = port_name
    )
  )

>>>>

# Include figure for paper

land_sf$group <- factor(state_mpa$abrev,      # Reordering group factor levels
                          levels = state_lat$abrev)

ggplot() +
  geom_sf(data = st_as_sf(map("state", plot = FALSE, fill = TRUE))) +
  geom_sf(data = land_sf, aes(fill = state)) +
  scale_fill_manual(values = state_pallet) +
  my_ggtheme_m(leg_pos = "right")

@@

state_sf <-  st_read("/Volumes/Enterprise/Data/Spatial/FederalAndStateWaters") %>%
  janitor::clean_names() %>% 
  mutate(jurisdicti = str_to_lower(jurisdicti)) %>% 
  filter(jurisdicti %in% States) %>% 
  mutate(state = str_to_sentence(jurisdicti)) %>% 
  st_transform(crs = 4326) %>%   # for matching projections
  st_simplify(preserveTopology = TRUE, dTolerance = 10000) #0.1 for paper

@@

# quantiles_df <- overall_port_data %>% 
#   group_by(species_name) %>% 
#   summarise(
#     "95th" = quantile(year_ton, 0.95),
#     "25th" = quantile(year_ton, 0.75),
#     "50th" = quantile(year_ton, 0.50)) %>% 
#   gather("tresh","quant",`95th`:`50th`)
# 
# 
# top_portsb <- overall_port_data %>% 
#   left_join(quantiles_df,
#             by = "species_name") %>% 
#   mutate(
#     treshold = ifelse(year_ton >= quant,tresh,NA) 
#   )

@@

# Visual exploration of all together (o.k! )
ggplot() +
  geom_sf(data = eez_sf, aes()) +
  geom_sf(data = land_sf, aes(fill = state), alpha = 0.3) +
  geom_sf(data = state_sf, aes(color = state), fill = NA) +
  geom_point(data = subset(noaa_ports_complete, tresh == "25th"), aes(x = lon, y = lat,
                                                                      shape = species_name), size = 3, alpha = 0.7) +
  scale_color_manual(values = state_pallet) +
  scale_fill_manual(values = state_pallet) +
  theme_dark() +
  coord_sf(ylim = c(33,48))

@@

#Load top port data 
ports_df <- my_path("G","AcrossBoundaries/Data/Partial", "top_ports.csv", read = T) %>% 
  filter(tresh == "25th")

@@

# Ports data
ports_df <- my_path("G","AcrossBoundaries/Data/Partial", "top_ports.csv", read = T) %>% 
  filter(tresh == "25th") %>%
  group_by(species_name,state_postal) %>% 
  summarise(n =n()) %>% 
  select(-n)
# distinct(c(,state_postal), .keep_all = T) # No need for duplication

@@

# length(state_pallet)

@@

## F0 Map of the US and study site

# re order states by latitude
land_sf_t <- land_sf

land_sf_t$group <- factor(land_sf_t$abrev,      # Reordering group factor levels
                               levels = state_lat$abrev)

 land_sf_t <- land_sf_t %>% 
  left_join(state_lat) %>% 
   mutate(
     name = paste0(str_to_title(state)," (",abrev,")")
   )

 # Load us MAP
 us_map <- st_as_sf(map("state", plot = FALSE, fill = TRUE))

 # Plot figure
study_plot <-
  ggplot() +
  geom_sf(data = us_map, aes(),fill = "grey", color = "lightgrey") +
  geom_sf(data = land_sf_t, aes(fill = reorder(name,-order))) +
  scale_fill_manual("State",
                    values = state_pallet) +
  my_ggtheme_m(leg_pos = "right",map_type = "R")

name_f0 <- my_path("G", "AcrossBoundaries/Results/Figures", name = "study_region.png")

ggsave(name_f0,
       study_plot,
       width = 9,
       height = 4)

@@

my_path("D","Partial","top_ports.csv", read = T) %>% 
  filter(tresh == "25th",
         state_postal == "NY") %>% 
  ggplot() +
  geom_point(
    aes(
      x = lon,
      y = lat,
      color = port_name
    )
  )
